---
title: "Chest x-ray survey analysis"
output: html_document
date: "2026-02-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(haven)
library(tidyverse)
library(ggplot2)
library(Kendall)
```

## Read survey data

```{r cars}
data <- read_sav('Survey data.sav')
```

## Fully completed surveys

```{r}
# determine number of fully completed surveys

data_closed <- select(data, PatientID, Age, Sex, FormalEducation, AILevelofKnowledge,
                      AIConfidence, CareType, Consent, PotentialConcerns, ImpactOnCare,
                      AIReliance, RiskStratification)


data_closed <- na.omit(data_closed)
nrow(na.omit(data_closed))

# calculate "complete" marker (1 if survey was fully completed, 0 if partially completed)
data <- data %>% 
  mutate(complete = if_else(PatientID %in% data_closed$PatientID, 1, 0))

# calculate whether there is significant difference in sex between participants who partially or fully completed the survey
chisq.test(data$Sex, data$complete)

# calculate whether there is significant difference in age between participants who partially or fully completed the survey
t.test(data$Age[data$complete == 1], y = data$Age[data$complete == 0])
```

## Sex

```{r}
# number of men and women in the dataset
table(data$Sex, useNA = "ifany")
```

## Age

```{r}
# age histogram
hist(data$Age)

# mean and standard deviation of age
mean(data$Age, na.rm = TRUE)
sd(data$Age, na.rm = TRUE)

# minimum and maximum age
min(data$Age, na.rm = TRUE)
max(data$Age, na.rm = TRUE)

# age categorised
data$Age_cat <- cut(data$Age, 
                   breaks=c(0, 19, 34, 49, 64, 79, 100))

# frequency table of age categories
table(data$Age_cat, useNA = "ifany")

# percentage frequency table of age categories
table(data$Age_cat, useNA = "ifany")/nrow(data)*100
```

## Educational Background

```{r}
# frequency and percentage tables of Educational Background

table(data$FormalEducation, useNA = "ifany")

table(data$FormalEducation, useNA = "ifany")/nrow(data)*100
```
# AI knowledge

```{r}
# Labels:
#  value           label
#      1         Nothing
#      2        A little
#      3 Moderate amount
#      4           A lot

table(data$AILevelofKnowledge, useNA = "ifany")

# amendment to ggplot to avoid code repetition
ggplot <- function(...) ggplot2::ggplot(...) +
  geom_col(position = 'dodge', fill = "#756bb1", alpha = .5, colour = "#756bb1") + 
  geom_text(position = position_dodge(width = .9),    # move to center of bars
            vjust = -0.5,    # nudge above top of bar
            size = 3.5) +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(size=14),
        axis.title.x = element_text(vjust=-1),
        axis.title.y = element_text(vjust=2),
        panel.grid.major = element_line(linewidth = 1),
        panel.grid.major.x = element_blank())

# barplot of self-reported AI knowledge
data %>% 
  filter(!is.na(AILevelofKnowledge)) %>% 
  count(AILevelofKnowledge = factor(AILevelofKnowledge)) %>% 
  mutate(pct = prop.table(n)) %>% 
  ggplot(aes(x = AILevelofKnowledge, y = n, label = scales::percent(pct))) + 
  labs(title = "How much do you know about the use of artificial intelligence in healthcare?",
       y = "Number of respondents", x = "Knowledge of artificial intelligence in healthcare") +
  scale_x_discrete(labels=c("Nothing",  "A little", "A moderate amount","A lot")) +
  coord_cartesian(clip = "off")
```
## How comfortable are you with the use of AI?

```{r}
table(data$AIConfidence, useNA = "ifany")

# barplot of comfort with AI in care
data %>% 
  filter(!is.na(AIConfidence)) %>% 
  count(AIConfidence = factor(AIConfidence)) %>% 
  mutate(pct = prop.table(n)) %>% 
  ggplot(aes(x = AIConfidence, y = n, label = scales::percent(pct))) + 
  labs(title = "How comfortable are you with the use of artificial intelligence technology \n in your care?",
       y = "Number of respondents", x = "Approval of artificial intelligence use") +
  scale_x_discrete(labels=c("Object", "Neutral", "Approve", "Strongly approve", "I don't know")) +
  coord_cartesian(clip = "off")
```

## How should AI be used?

```{r}
# Labels:
#  value          label
#      1       AI Alone
#      2 A Decision Aid
#      3     Not at all
#      4    No response

table(data$CareType, useNA = "ifany")
```

# Opportunity to opt-out

```{r}
# Labels:
#  value         label
#      1           Yes
#      2            No
#      3 I do not know
#      4   No Response
#      

table(data$Consent, useNA = "ifany")
```

# Concerns

```{r}
# Labels:
#  value                         label
#      1     AI tool may make mistakes
#      2       Privacy and/or security
#      3                   No concerns
#      4 Both Mistake & Privacy Issues
#      5                   No Response
     
table(data$PotentialConcerns, useNA = "ifany")
```

# Would it affect your care?

```{r}
# Labels:
#  value         label
#      1   Worsen care
#      2     No change
#      3  Improve care
#      4 I do not know
#      5   No response

table(data$ImpactOnCare, useNA = "ifany")
```

## Will doctors become reliant?

```{r}
# Labels:
#  value         label
#      1           YES
#      2         Maybe
#      3            NO
#      4 I do not know
#      5   No response

table(data$AIReliance, useNA = "ifany")
```

## Approval of AI risk stratification

```{r}
table(data$RiskStratification, useNA = "ifany")

# barplot of AI risk stratification approval
data %>% 
  filter(!is.na(RiskStratification)) %>% 
  count(RiskStratification = factor(RiskStratification)) %>% 
  mutate(pct = prop.table(n)) %>% 
  ggplot(aes(x = RiskStratification, y = n, label = scales::percent(pct))) + 
  labs(title = "This artificial intelligence tool determines the order in which chest X-rays are \n seen by doctors, in order to prioritise patients who need urgent clinical care. \n To what extent do you approve? ",
       y = "Number of respondents", x = "Approval of artificial intelligence use for prioritisation") +
  scale_x_discrete(labels=c("Strongly disapprove", "Disapprove", "Neutral", "Approve", "Strongly approve")) +
  coord_cartesian(clip = "off")
```

## Correlation between participants' AI knowledge & AI acceptance

```{r}
# Labels:
#  value            label
#      1  Strongly object
#      2           Object
#      3          Neutral
#      4          Approve
#      5 Strongly approve
#      6     I don't know

# exclude missing or I don't know responses
data_corr <- data %>% 
  filter(!is.na(data$AILevelofKnowledge),
         !is.na(data$AIConfidence),
         data$AIConfidence != 6)

summary(Kendall(data_corr$AILevelofKnowledge, data_corr$AIConfidence))
```
### 95% CI

```{r}
calc_tau_boot = function(data, indices){
  
  # this function produces a statistic to be bootstrapped
  # it returns the following statistic: Kendall's tau-b

  # expected input: a dataframe (data) with two columns
  
  # the function is used as an input for boot() from the 'boot' package
  
  dt <- data[indices,]
  var_1 <- dt[,1]
  var_2 <- dt[,2]
  
  output <- Kendall(var_1, var_2)
  
  return(output$tau)
}


# calculate 95% confidence interval (CI) with bootstrapping for Kendall tau-b

set.seed(10)

reps <- boot(data.frame(data_corr$AILevelofKnowledge, data_corr$AIConfidence), calc_tau_boot, R = 5000)

# calculate the CIs
tau_ci <- boot.ci(reps, conf=0.95, type = c("bca"))

tau_ci$bca[4:5]
```

## Correlation between participants' AI knowledge & approval of AI prioritisation

```{r}
# Labels:
#  value               label
#      1 Strongly disapprove
#      2          Disapprove
#      3             Neutral
#      4             Approve
#      5    Strongly Approve
#      6         No response

# exclude missing or I don't know responses
data_corr <- data %>% 
  filter(!is.na(data$AILevelofKnowledge),
         !is.na(data$RiskStratification),
         data$RiskStratification != 6)

summary(Kendall(data_corr$AILevelofKnowledge, data_corr$RiskStratification))
```
### 95% CI

```{r}
# calculate 95% confidence interval (CI) with bootstrapping for Kendall tau-b

reps <- boot(data.frame(data_corr$AILevelofKnowledge, data_corr$RiskStratification), calc_tau_boot, R = 5000)

# calculate the CIs
tau_ci <- boot.ci(reps, conf=0.95, type = c("bca"))

tau_ci$bca[4:5]
```
